---
title: "prelim_mcmc"
author: "Nathan Welch"
date: "3/29/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, cache=FALSE}
library(ggplot2)
library(xtable)
library(denstrip)
library(dplyr)
library(MASS)
library(RColorBrewer)
library(gridExtra)
library(gridBase)
library(grid)
library(raster)
library(coda)
```

# Data Cleanup

```{r}
# Data Cleanup
df = read.table("/Users/nwelch/prelim/data/canedata.txt")

names(df) = c("x","y", paste0("t", 1:6))

df[df$x==10.2,"x"] = 10.5

everInfected = as.logical(df$t6)

dfNeverInfected = df[!everInfected,]
tmp = df[everInfected,]

theorder = order(tmp$t1, tmp$t2, tmp$t3, tmp$t4, tmp$t5, tmp$t6, decreasing=TRUE)

dfInfected = tmp[theorder,]

df = rbind(dfNeverInfected, dfInfected)
rownames(df) = seq(1, nrow(df), 1)

# Infection Metrics
cumInfected = c(0, apply(df[,-(1:2)], 2, sum))
timeInspected = c(0, 6, 10, 14, 19, 23, 30)
inspectionGap = c(0, diff(timeInspected))

midInspection = (timeInspected - inspectionGap/2)[-1]
infectedIndicator = dfInfected[dfInfected$t6==1, 3:8]
tmp = mapply('*', infectedIndicator, midInspection)

notZeroFunc = function(r){
  notZero = r[r>0] 
  min(notZero)
}

tau = apply(tmp, 1, notZeroFunc)

tauUpperBoundTmp = mapply('*', infectedIndicator, timeInspected[-1])
tauUpperBound = apply(tauUpperBoundTmp, 1, notZeroFunc)

whichZeroFunc = function(r){
  notZero = r[r>0] 
  minNotZero = min(notZero)
  which(r==minNotZero)
}

tauLowerBoundIndex = apply(tauUpperBoundTmp, 1, whichZeroFunc)
getMinTime = function(i) timeInspected[i]
tauLowerBound = sapply(tauLowerBoundIndex, getMinTime)

tmp = cbind(dfNeverInfected, 40, 40, 40)
infectionData = cbind(dfInfected, tau, tauLowerBound, tauUpperBound)
names(tmp) = names(infectionData)
df = rbind(tmp, infectionData)
write.csv(df, file="/Users/nwelch/prelim/data/plantData.csv", row.names=FALSE)

xi = df$x + 1.0i*df$y
dxixj = Mod(outer(xi, xi, FUN="-"))
write.csv(dxixj, file="/Users/nwelch/prelim/data/dxixj.csv", 
          row.names=FALSE)
```

# Create Test Data Set
The following codes create a test set that is a small sample of the data from the field. The ideas is that this data offers a reasonably computable sample that should lead us to somewhat reasonable estimates of the full data set parameter values. 
```{r}
reLimit = max(Re(xi))*sqrt(0.1)
imLimit = max(Im(xi))*sqrt(0.1)
xiSample = xi[Re(xi)<=reLimit & Im(xi)<=imLimit]
testPlants = which((df$x <= reLimit) & (df$y <= imLimit))
dfTest = df[testPlants,]

write.csv(dfTest, file="/Users/nwelch/prelim/data/plantDataTest.csv",
          row.names=FALSE)

dxixjSample = Mod(outer(xiSample, xiSample, FUN="-"))
write.csv(dxixjSample, file="/Users/nwelch/prelim/data/dxixjTest.csv",
          row.names=FALSE)
```

# Figure Reproduction

```{r, echo=FALSE}
fig1a = ggplot(data=df, aes(x, y, color=factor(t6)), legend=FALSE) + 
  geom_point() + 
  scale_colour_manual(values = c("0"="grey", "1"="black"), guide=FALSE) +
  theme_bw() +
  theme(axis.line = element_line(colour='black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    plot.margin = unit(c(1, 0.5, 0.5, 0), "lines"),
    axis.text=element_text(size=12),
    axis.title=element_text(size=12))+
  xlab("meters")+
  ylab("meters") + 
  geom_rect(aes(xmin=1, xmax=round(reLimit), ymin=-1, ymax=round(imLimit)),
            fill="transparent", color="blue", size=1, lty=2)
ggsave(filename="/Users/nwelch/prelim/report/figures/figure_1a.png", fig1a, 
       dpi=300, height=4.5, width=4)
fig1a
```

```{r, echo=FALSE}

infectionDf = data.frame(cumInfected, inspectionGap)
rownames(infectionDf)[1]="t0"

tmp = apply(infectionDf, 1, function(x) rep(x[1], x[2]))
infectedSeries = unlist(tmp, use.names=FALSE)

week = seq(1, length(infectedSeries), 1)

expandedInfectionDf = data.frame(week, infectedSeries)
```

```{r, echo=FALSE}
fig1b = ggplot(data=expandedInfectionDf, 
                aes(week, infectedSeries), legend=FALSE) + 
  geom_step() +
  theme_bw() +
  theme(axis.line = element_line(colour='black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(), 
    plot.margin = unit(c(1, 0.5, 0.5, 0), "lines"),
    axis.text=element_text(size=12),
    axis.title=element_text(size=12))+
  xlab("Weeks")+
  ylab("Number infected")
ggsave(filename="/Users/nwelch/prelim/report/figures/figure_1b.png", fig1b, 
       dpi=300, height=4.5, width=4)
fig1b
```

```{r}
g = grid.arrange(fig1a, fig1b, ncol=2)
g
ggsave(filename="/Users/nwelch/prelim/report/figures/figure_1.png", g, 
       dpi=300, height=4, width=6)
```
# Implementation Performance
```{r results = 'asis', echo=FALSE}
theta = c(28.15, 25.08, 2.38, 1.07, 261.58, 16.65)
mu = c(14.27, 12.57, 0.25, 1.07, 253.65, 0.35)
tau = c(160.92, 29.99, 10.18, 1.76, 122.81, 2.15)
sigma = c(28.99, 25.17, 12.31, 1.13, 263.19, 16.80)

tmp = data.frame(theta, mu, tau, sigma)
rownames(tmp) = c('Basic', 'Parallel', 'Improved', 'Truncated', 
                'Basic Discrete', 'Truncated Discrete')

performance = as.data.frame(cbind(tmp, rowSums(tmp)))
names(performance)[5]='Total'

addtorow <- list()
addtorow$pos <- list(-1)
addtorow$command <- c("Algorithm & \\multicolumn{5}{c}{Times for Updating the Following Parameters} \\\\\n") 

tmp = performance

colnames(tmp) = c("$\\theta$", "$\\mu$", "$\\tau$", "$\\sigma$", "Total")
tmp = xtable(tmp, caption = "Table with xtable")

print(tmp, add.to.row = addtorow, hline.after=c(-1, 0, nrow(tmp)),
      sanitize.colnames.function = function(x) {x})
```
## MCMC Chain Evaluation
```{r}
fullChain = read.csv(file="~/prelim/data/mcmc_chain_julia.csv")
results = fullChain[10000:nrow(fullChain),]
accept_results = read.csv(file="~/prelim/data/mcmc_chain_metrics_julia.csv")
number_of_trials = nrow(results)
accept_results
```

```{r}
p = ggplot(results) + 
  geom_path(aes(x=1:number_of_trials, y=mu)) +
  theme_bw() +
  theme(axis.text=element_text(size=12), axis.title=element_text(size=12)) +
  ylab(expression(mu)) +
  xlab("Metropolis Iteration After Burn-in")
p
ggsave(filename="/Users/nwelch/prelim/report/figures/mu_mcmc_chain.png", p, 
       dpi=300, height=4, width=6)
```

```{r}
p = ggplot(results) + 
  geom_path(aes(x=1:number_of_trials, y=theta)) +
  theme_bw() +
  theme(axis.text=element_text(size=12), axis.title=element_text(size=12)) +
  ylab(expression(theta)) +
  xlab("Metropolis Iteration After Burn-in")
p
ggsave(filename="/Users/nwelch/prelim/report/figures/theta_mcmc_chain.png", p, 
       dpi=300, height=4, width=6)
```

```{r}
p = ggplot(results) + 
  geom_path(aes(x=1:number_of_trials, y=sigma)) +
  theme_bw() +
  theme(axis.text=element_text(size=12), axis.title=element_text(size=12)) +
  ylab(expression(sigma)) +
  xlab("Metropolis Iteration After Burn-in")
p
ggsave(filename="/Users/nwelch/prelim/report/figures/sigma_mcmc_chain.png", p, 
       dpi=300, height=4, width=6)
```
### Convergence Tests
```{r}
MCMC.chain = mcmc(data=results[,c('mu', 'sigma', 'theta')], 
                  start = 1, end = nrow(results), thin = 1)
MCMC.mu = mcmc(data=results[,'mu'], start = 1, end = nrow(results), thin = 1)
MCMC.sigma = mcmc(data=results[,'sigma'], start = 1, end = nrow(results), thin = 1)
MCMC.theta = mcmc(data=results[,'theta'], start = 1, end = nrow(results), thin = 1)
```

```{r}
chainTail = (number_of_trials-1e4):number_of_trials
png(filename="/Users/nwelch/prelim/report/figures/mcmc_cum_quant_plot_mu.png", 
    res=300, width=4.5, height=4, units='in')
par(mar=c(3,3,1,0.5), mgp=c(1.5,0.5,0))
cumuplot(MCMC.mu[chainTail], probs=c(0.025,0.5,0.975), auto.layout = F,
         main=expression(paste("Cumulative Quantile Plot, ", mu)), 
         ylab=expression(paste(mu, " 0.025, 0.5, and 0.975 quantile")),
         xlab="Last 10,000 Metropolis Iterations")
par(mar=c(5,4,4,1)+0.1, mgp=c(3,1,0))
dev.off()
```
```{r}
png(filename="/Users/nwelch/prelim/report/figures/mcmc_cum_quant_plot_sigma.png", 
    res=300, width=4.5, height=4, units='in')
par(mar=c(3,3,1,0.5), mgp=c(1.5,0.5,0))
cumuplot(MCMC.sigma[chainTail], probs=c(0.025,0.5,0.975), auto.layout = F,
         main=expression(paste("Cumulative Quantile Plot, ", sigma)), 
         ylab=expression(paste(sigma, " 0.025, 0.5, and 0.975 quantile")),
         xlab="Last 10,000 Metropolis Iterations")
par(mar=c(5,4,4,1)+0.1, mgp=c(3,1,0))
dev.off()
```
```{r}
png(filename="/Users/nwelch/prelim/report/figures/mcmc_cum_quant_plot_theta.png", 
    res=300, width=4.5, height=4, units='in')
par(mar=c(3,3,1,0.5), mgp=c(1.5,0.5,0))
cumuplot(MCMC.theta[chainTail], probs=c(0.025,0.5,0.975), auto.layout = F,
         main=expression(paste("Cumulative Quantile Plot, ", theta)), 
         ylab=expression(paste(theta, " 0.025, 0.5, and 0.975 quantile")),
         xlab="Last 10,000 Metropolis Iterations")
par(mar=c(5,4,4,1)+0.1, mgp=c(3,1,0))
dev.off()
```

```{r}
png(filename="/Users/nwelch/prelim/report/figures/mcmc_acf_plot_mu.png", 
    res=300, width=4.5, height=4, units='in')
par(mar=c(3,3,1,0.5), mgp=c(1.5,0.5,0))
autocorr.plot(MCMC.mu, lag.max=30,
              main=expression(paste("Post Burn-in ACF Plot, ", mu)))
par(mar=c(5,4,4,1)+0.1, mgp=c(3,1,0))
dev.off()
```
```{r}
png(filename="/Users/nwelch/prelim/report/figures/mcmc_acf_plot_sigma.png", 
    res=300, width=4.5, height=4, units='in')
par(mar=c(3,3,1,0.5), mgp=c(1.5,0.5,0))
autocorr.plot(MCMC.sigma, lag.max=30, 
              main=expression(paste("Post Burn-in ACF Plot, ", sigma)))
par(mar=c(5,4,4,1)+0.1, mgp=c(3,1,0))
dev.off()
```
```{r}
png(filename="/Users/nwelch/prelim/report/figures/mcmc_acf_plot_theta.png", 
    res=300, width=4.5, height=4, units='in')
par(mar=c(3,3,1,0.5), mgp=c(1.5,0.5,0))
autocorr.plot(MCMC.theta, lag.max=30, 
              main=expression(paste("Post Burn-in ACF Plot, ", theta)))
par(mar=c(5,4,4,1)+0.1, mgp=c(3,1,0))
dev.off()
```

```{r}
effectiveSize(MCMC.chain)
```

```{r}
geweke.diag(MCMC.chain, frac1=0.1, frac2=0.5)
```

```{r}
raftery.diag(MCMC.chain, q=0.95, r=0.01, s=0.95, converge.eps=0.001)
```

# Parameter Estimation
```{r}
thin = round(number_of_trials / min(effectiveSize(MCMC.chain)))
thinResults = results[seq(1, number_of_trials, thin),]
write.csv(thinResults, file="/Users/nwelch/prelim/data/thinResults.csv",
          row.names=FALSE)
```

```{r}
ss = 1e4
rate = rgamma(ss, shape=0.7, rate=0.004)
scale = rgamma(ss, shape=0.7, scale=0.004)
rateScale=data.frame(rate, scale)


p = ggplot(data=rateScale)+
    theme(axis.line = element_line(colour='black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank())

prate = p+xlim(0, 800)+geom_density(aes(x=rate))
pscale = p+xlim(0, 0.01)+geom_density(aes(x=scale))

g = grid.arrange(prate, pscale, ncol=2)
g
ggsave(filename="/Users/nwelch/prelim/report/figures/rateScaleCompare.png", g, 
       dpi=300, height=4, width=6)
```

```{r, echo=FALSE}
prior_mu = rgamma(10000, shape=1.307877, rate=0.06334962) #, mu_rate=
post_mu = thinResults$mu

par(mar=c(4,5,1,1))
stripwidth=0.5
plot(NA, xlim=c(0, 80), ylim=c(0,2),
		xlab=expression(mu), yaxt="n", ylab="")
axis(2, at=c(0.5, 1.5), labels=c("Posterior", "Prior"), las=1)

denstrip(prior_mu, horiz=T, at=1.5, 
		ticks=quantile(prior_mu, probs=c(0.025, 0.5, 0.975)), width=stripwidth)

denstrip(post_mu, horiz=T, at=0.5,
		ticks=quantile(post_mu, probs=c(0.025, 0.5, 0.975)), width=stripwidth)
par(mar=c(5,4,4,2)+0.1)
```
```{r, echo=FALSE}
xseq = seq(0, 1, len=5e4)[-(1:4)]
stripwidth=0.5

png(filename="/Users/nwelch/prelim/report/figures/rate_v_scale_plot.png", 
    res=300, width=4, height=2, units='in')
par(mar=c(3,5,1,0.5), mgp=c(2,0.5,0.5))

plot(NA, xlim=c(0, 0.008), ylim=c(0,3),
		xlab=expression(mu), yaxt="n", ylab="")
axis(2, at=c(2.5, 1.5, 0.5), labels=c("Error", "Rate", "Scale"), las=1)

# figure error
denstrip(xseq, dgamma(xseq, shape=0.7, scale=0.7), horiz=T, at=2.5, width=stripwidth)

# rate plot
denstrip(xseq, dgamma(xseq, shape=0.7, rate=0.004), horiz=T, at=1.5, 
		ticks=qgamma(c(0.025, 0.5, 0.975), shape=0.7, rate=0.004), width=stripwidth)

# scale error
gammaMedian=NULL #qgamma(0.5, shape=0.7, scale=0.004)
denstrip(xseq, dgamma(xseq, shape=0.7, scale=0.004), horiz=T, at=0.5,
		ticks=c(0.7*0.004, gammaMedian), width=stripwidth)

par(mar=c(5,4,4,2)+0.1, mgp=c(3,1,0))
dev.off()
```
```{r, echo=FALSE}
prior_mu = rgamma(10000, shape=1.307877, rate=0.06334962) #, mu_rate=
post_mu = thinResults$mu

png(filename="/Users/nwelch/prelim/report/figures/figure_2a.png", 
    res=300, width=4, height=2, units='in')
par(mar=c(2.35,5,1,0.5), mgp=c(1.5,0.5,0), cex.axis=1.25, cex.lab=1.25)
stripwidth=0.5
plot(NA, xlim=c(0, 0.01), ylim=c(0,2),
		xlab=expression(mu), yaxt="n", ylab="")
axis(2, at=c(0.5, 1.5), labels=c("Posterior", "Prior"), las=1)

denstrip(prior_mu, horiz=T, at=1.5, 
		ticks=quantile(prior_mu, probs=c(0.025, 0.5, 0.975)), width=stripwidth)

denstrip(post_mu, horiz=T, at=0.5,
		ticks=quantile(post_mu, probs=c(0.025, 0.5, 0.975)), width=stripwidth)
par(mar=c(5,4,4,2)+0.1, mgp=c(3,1,0), cex.axis=1, cex.lab=1)
dev.off()
```

```{r, echo=FALSE}
prior_sigma = rgamma(10000, shape=0.5, scale=100)
post_sigma = thinResults$sigma

par(mar=c(4,5,1,1))
stripwidth=0.5
plot(NA, xlim=c(0, 250), ylim=c(0,2),
		xlab=expression(sigma), yaxt="n", ylab="")
axis(2, at=c(0.5, 1.5), labels=c("Posterior", "Prior"), las=1)

denstrip(prior_sigma, horiz=T, at=1.5, 
		ticks=quantile(prior_sigma, probs=c(0.025, 0.5, 0.975)), width=stripwidth)

denstrip(post_sigma, horiz=T, at=0.5,
		ticks=quantile(post_sigma, probs=c(0.025, 0.5, 0.975)), width=stripwidth)
par(mar=c(5,4,4,2)+0.1)
```

```{r, echo=FALSE}
prior_sigma = rgamma(10000, shape=0.5, scale=100)
post_sigma = thinResults$sigma

png(filename="/Users/nwelch/prelim/report/figures/figure_2c.png", 
    res=300, width=4, height=2, units='in')
par(mar=c(2.35,5,1,0.5), mgp=c(1.5,0.5,0), cex.axis=1.25, cex.lab=1.25)
stripwidth=0.5
plot(NA, xlim=c(0.5, 2.5), ylim=c(0,2),
		xlab=expression(sigma), yaxt="n", ylab="")
axis(2, at=c(0.5, 1.5), labels=c("Posterior", "Prior"), las=1)

denstrip(prior_sigma, horiz=T, at=1.5, 
		ticks=quantile(prior_sigma, probs=c(0.025, 0.5, 0.975)), width=stripwidth)

denstrip(post_sigma, horiz=T, at=0.5,
		ticks=quantile(post_sigma, probs=c(0.025, 0.5, 0.975)), width=stripwidth)
par(mar=c(5,4,4,2)+0.1, mgp=c(3,1,0), cex.axis=1, cex.lab=1)
dev.off()
```

```{r, echo=FALSE}
prior_theta = rgamma(10000, shape=0.8, rate=0.1)
post_theta = thinResults$theta

par(mar=c(4,5,1,1))
stripwidth=0.5
plot(NA, xlim=c(0, 35), ylim=c(0,2),
		xlab=expression(theta), yaxt="n", ylab="")
axis(2, at=c(0.5, 1.5), labels=c("Posterior", "Prior"), las=1)

denstrip(prior_theta, horiz=T, at=1.5, 
		ticks=quantile(prior_theta, probs=c(0.025, 0.5, 0.975)), width=stripwidth)

denstrip(post_theta, horiz=T, at=0.5,
		ticks=quantile(post_theta, probs=c(0.025, 0.5, 0.975)), width=stripwidth)
par(mar=c(5,4,4,2)+0.1)
```
```{r, echo=FALSE}
prior_theta = rgamma(10000, shape=0.8, rate=0.1)
post_theta = thinResults$theta

png(filename="/Users/nwelch/prelim/report/figures/figure_2b.png", 
    res=300, width=4, height=2, units='in')
par(mar=c(2.35,5,1,0.5), mgp=c(1.5,0.5,0), cex.axis=1.25, cex.lab=1.25)
stripwidth=0.5
plot(NA, xlim=c(0.05, 0.2), ylim=c(0,2),
		xlab=expression(theta), yaxt="n", ylab="")
axis(2, at=c(0.5, 1.5), labels=c("Posterior", "Prior"), las=1)

denstrip(prior_theta, horiz=T, at=1.5, 
		ticks=quantile(prior_theta, probs=c(0.025, 0.5, 0.975)), width=stripwidth)

denstrip(post_theta, horiz=T, at=0.5,
		ticks=quantile(post_theta, probs=c(0.025, 0.5, 0.975)), width=stripwidth)
par(mar=c(5,4,4,2)+0.1, mgp=c(3,1,0), cex.axis=1, cex.lab=1)
dev.off()
```

```{r}
ci = sort(c(0.5, 0.8, 0.95), decreasing=TRUE)

ci_contours = function(x, y) {
  kernelHat = kde2d(x, y, n=200)
  kernelHatPdf = kernelHat$z
  ordering = order(kernelHatPdf, decreasing=TRUE)
  sumKernelHat = cumsum(kernelHatPdf[ordering])
  cdfHat = sumKernelHat/max(sumKernelHat)
  
  getCdfCutoff = function(a) kernelHatPdf[ ordering[ max( which( cdfHat < a) ) ] ]
  lvls = sapply(ci, getCdfCutoff)
  
  contourList = contourLines(kernelHat, levels=lvls)
  contourDataFrame = lapply(contourList, data.frame)
  
  return(contourDataFrame)
}

```

```{r, echo=FALSE}
contours = ci_contours(thinResults$sigma, thinResults$mu)

fig2d = ggplot(data=thinResults, aes(sigma, mu)) + 
  geom_point(color='grey') +
  geom_path(data=contours[[1]], aes(x, y), color="grey20", size=1) +
  geom_path(data=contours[[2]], aes(x, y), color="grey40", size=1) +
  geom_path(data=contours[[3]], aes(x, y), color="grey60", size=1) +
  theme(axis.line = element_line(color='black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        plot.margin = unit(c(1, 0.5, 0.5, 0.75), "lines"),
        axis.text=element_text(size=16),
        axis.title=element_text(size=16))+
  xlab(expression(sigma))+
  ylab(expression(mu))+
  xlim(0.5,2.5)
ggsave(filename="/Users/nwelch/prelim/report/figures/figure_2d.png", fig2d, 
       dpi=300, height=4, width=4)
fig2d
```
```{r, echo=FALSE}
contours = ci_contours(thinResults$theta, thinResults$mu)

fig2e = ggplot(data=thinResults, aes(theta, mu)) + 
  geom_point(color='grey') +
  geom_path(data=contours[[1]], aes(x, y), color="grey20", size=1) +
  geom_path(data=contours[[2]], aes(x, y), color="grey40", size=1) +
  geom_path(data=contours[[3]], aes(x, y), color="grey60", size=1) +
  theme(axis.line = element_line(color='black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        plot.margin = unit(c(1, 0.5, 0.5, 0.75), "lines"),
        axis.text=element_text(size=16),
        axis.title=element_text(size=16))+
  xlab(expression(theta))+
  ylab(expression(mu))+
  xlim(0.05,0.2)
ggsave(filename="/Users/nwelch/prelim/report/figures/figure_2e.png", fig2e, 
       dpi=300, height=4, width=4)
fig2e
```
```{r, echo=FALSE}
contours = ci_contours(thinResults$theta, thinResults$sigma)

fig2f = ggplot(data=thinResults, aes(theta, sigma)) + 
  geom_point(color='grey') +
  geom_path(data=contours[[1]], aes(x, y), color="grey20", size=1) +
  geom_path(data=contours[[2]], aes(x, y), color="grey40", size=1) +
  geom_path(data=contours[[3]], aes(x, y), color="grey60", size=1) +
  theme(axis.line = element_line(color='black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        plot.margin = unit(c(1, 0.5, 0.5, 3), "lines"),
        axis.text=element_text(size=16),
        axis.title=element_text(size=16))+
  xlab(expression(theta))+
  ylab(expression(sigma))+
  ylim(0.5, 3)+
  xlim(0.05, 0.2)
ggsave(filename="/Users/nwelch/prelim/report/figures/figure_2f.png", fig2f, 
       dpi=300, height=4, width=4)
fig2f
```
# Simulated Epidemics 
```{r}
source("~/prelim/src/R/simulation.R")
colMeans(thinResults[,c('mu', 'sigma', 'theta')])
oneSim = scalarEpiSim(mu=0.004352591,
                      sigma=1.196941925,
                      theta=0.117329258,
                      simCount=1000)
hist(colSums(oneSim[,-1]), breaks=30)
timeSim = scalarEpiSim(returnSummary = FALSE)


u = c(0.0028, 0.0025, 0.003)
s = c(1.1, 1.15, 1.08)
t = c(0.11, 0.08, 0.15)
vecSim = vectorEpiSim(muSigmaThetaVector=cbind(u,s,t), simCount=100)
```

## Conditional Posterior Simulated Infections Per Week
```{r}
sampleTau = thinResults[,1:(ncol(thinResults)-3)]
pstSim = postEpiSim(sampleTau)

png(filename="/Users/nwelch/prelim/report/figures/figure_4a.png", 
    res=300, width=4.5, height=4, units='in')
par(mar=c(3.1, 3.1, 1, 1), mgp=c(2,0.5,0))
matplot(x=pstSim['weeks',], y=t(pstSim[seq(5, nrow(pstSim), 10),]), 
        type='l', lty=3, col="grey",
        ylab="infections per week", xlab="week")
matlines(x=pstSim['weeks',], y=t(pstSim[2:4,]), col='black', lty=c(1,2,2), lwd=3)
abline(v=timeInspected, lty=3, col="grey")
par(mfcol=c(1,1), mar=c(5,4,4,2)+0.1, mgp=c(3,1,0))
dev.off()
```

```{r}
sampleParams = thinResults[seq(1, nrow(thinResults), 100), c('mu', 'sigma', 'theta')]
vecSimSample = vectorEpiSim(muSigmaThetaVector=sampleParams, 
                            dxixj=dxixjSample, simCount=10)
```

```{r}
png(filename="/Users/nwelch/prelim/report/figures/figure_4b.png", 
    res=300, width=4.5, height=4, units='in')
par(mar=c(3.1, 3.1, 1, 1), mgp=c(2,0.5,0))
matplot(x=vecSimSample$week, y=vecSimSample[,5:ncol(vecSim)], type='l', 
        lty=3, col="grey", ylab="infections per week", xlab="week", ylim=c(0,10))
matlines(x=c(0, vecSimSample$week), y=rbind(c(0,0,0), vecSimSample[,2:4]),
         col='black', lty=c(1,2,2), lwd=3)
par(mfcol=c(1,1), mar=c(5,4,4,2)+0.1, mgp=c(3,1,0))
dev.off()
```

## All Plants Unconditional Simulation
```{r}
sampleParams = thinResults[seq(1, nrow(thinResults), 10), c('mu', 'sigma', 'theta')]
vecSim = vectorEpiSim(muSigmaThetaVector=sampleParams, dxixj=dxixj)
```

```{r}
png(filename="/Users/nwelch/prelim/report/figures/figure_4b_all_plants.png", 
    res=300, width=4.5, height=4, units='in')
par(mar=c(3.1, 3.1, 1, 1), mgp=c(2,0.5,0))
matplot(x=vecSim$week, y=vecSim[,5:ncol(vecSim)], type='l', lty=3, col="grey",
        ylab="infections per week", xlab="week", ylim=c(0,80))
matlines(x=c(0, vecSim$week), y=rbind(c(0,0,0), vecSim[,2:4]), 
         col='black', lty=c(1,2,2), lwd=3)
par(mfcol=c(1,1), mar=c(5,4,4,2)+0.1, mgp=c(3,1,0))
dev.off()
```

```{r}
png(filename="/Users/nwelch/prelim/report/figures/figure_4b_all_plants_root.png", 
    res=300, width=4.5, height=4, units='in')
par(mar=c(3.1, 3.1, 1, 1), mgp=c(2,0.5,0))
matplot(x=vecSim$week, y=sqrt(vecSim[,5:ncol(vecSim)]), type='l', lty=3, 
        col="grey", ylab="infections per week", xlab="week",
        ylim=sqrt(c(0,80)), yaxt="n")
theYseq = c(0,5,10,20,40,60,80)
axis(2, at=sqrt(theYseq), labels=theYseq)

matlines(x=c(0,vecSim$week), y=rbind(c(0,0,0), sqrt(vecSim[,2:4])), col='black', 
         lty=c(1,2,2), lwd=3)
par(mfcol=c(1,1), mar=c(5,4,4,2)+0.1, mgp=c(3,1,0))
dev.off()
```
## Density Plots
```{r}
tauPosterior = thinResults[,-((ncol(thinResults)-2):ncol(thinResults))]
firstKeep = colSums(tauPosterior <= round(min(apply(tauPosterior, 2, max))))>0
firstInfectionPosterior = tauPosterior[,firstKeep]
fromTo = round(c(min(firstInfectionPosterior), max(firstInfectionPosterior)))
breakPoints = seq(fromTo[1], fromTo[2],len=10)
tmp = apply(tauPosterior[,firstKeep, drop=F], 2, hist, breaks=breakPoints, plot=F)
postSampleDensity = NULL
for(psd in 1:length(tmp))
  postSampleDensity = cbind(postSampleDensity, 
                            tmp[[psd]]$density/sum(tmp[[psd]]$density))

### First time period
png(filename="/Users/nwelch/prelim/report/figures/figure_3a.png", 
    res=300, width=4.5, height=4, units='in')
par(mar=c(3.1, 3.1, 1, 1), mgp=c(2,0.5,0))
matplot(breakPoints[-1]-diff(breakPoints)/2, postSampleDensity, 
        type="l", lty=1, col="grey50",
        xlab="week", ylab="density", ylim=c(0, 0.2))
dev.off()
```

```{r}
lastKeep = colSums(tauPosterior >= round(max(apply(tauPosterior, 2, min))))>0
lastInfectionPosterior = tauPosterior[,lastKeep]
fromTo = round(c(min(lastInfectionPosterior), max(lastInfectionPosterior)))
breakPoints = seq(fromTo[1], fromTo[2],len=10)
tmp = apply(tauPosterior[,lastKeep, drop=F], 2, hist, breaks=breakPoints, plot=F)

### Last time period
postSampleDensity = NULL
for(psd in 1:length(tmp))
  postSampleDensity = cbind(postSampleDensity, 
                            tmp[[psd]]$density/sum(tmp[[psd]]$density))

postSampleMean = apply(postSampleDensity, 2, mean)
postSampleMean = matrix(postSampleMean, nrow(postSampleDensity), 
                        ncol(postSampleDensity))
centering = postSampleDensity - postSampleMean
centeringL = apply(centering[1:3,], 2, sum)
centeringH = apply(centering[7:9,], 2, sum)

png(filename="/Users/nwelch/prelim/report/figures/figure_3b.png", 
    res=300, width=4.5, height=4, units='in')
par(mar=c(3.1, 3.1, 1, 1), mgp=c(2,0.5,0))
plotLines=c(order(centeringL, decreasing=T)[c(1,10)],
            order(centeringH, decreasing=T)[c(1,10)])

matplot(breakPoints[-1]-diff(breakPoints)/2, postSampleDensity, 
        type="l", lty=1, col="grey80", xlab="week", ylab="density", 
        ylim=c(0, 0.2))
matlines(breakPoints[-1]-diff(breakPoints)/2, 
         postSampleDensity[,plotLines], 
         col="black",lty=c(1,1,2,2), lwd=1.5)
par(mfcol=c(1,1), mar=c(5,4,4,2)+0.1, mgp=c(3,1,0))
dev.off()
```
# Post 30 Weeks Prediction 
```{r}
df = read.csv(file="/Users/nwelch/prelim/data/plantData.csv")
susceptable = ifelse(df$t6==0, TRUE, FALSE)
onePredSim = scalarEpiSim(startTime=30, endTime=Inf, dxixj=dxixj, 
                          susceptableAtStart=susceptable,
                          returnSummary=FALSE)
sampleParams = thinResults[seq(1, nrow(thinResults), 100), c('mu', 'sigma', 'theta')]
manyPredSim = vectorEpiSim(muSigmaThetaVector=sampleParams,
                           startTime=30, endTime=Inf, dxixj=dxixj,
                           susceptableAtStart=susceptable,
                           returnSummary=FALSE)

cumTimes = c(35, 40, 50, 60)
colBreaks = c(-0.1, 0.2, 0.8, 0.9999, 1.01)
colLabel = c('zeroTo19', 'twentyTo79', 'eightyTo99', 'knownInfection')
probInfected = probInfLabel = NULL
for(ct in cumTimes){
  probabilities = apply(manyPredSim, 1, function(t) mean(t<ct))
  probInfected = cbind(probInfected, probabilities)
  probLabels = as.character(cut(probabilities, breaks=colBreaks, labels=colLabel))
  probInfLabel = cbind(probInfLabel, probLabels)
}
colnames(probInfected) = colnames(probInfLabel) = paste0("week", cumTimes)

prediction = data.frame(df[,c("x", "y")], fig5Labels)
```

```{r}
colGrey = grey(c(1, 0.65, 0.35, 0.05))
colLabel = c('zeroTo19', 'twentyTo79', 'eightyTo99', 'knownInfection')
thecol = c("zeroTo19"="grey100", "twentyTo79"="grey80", 
           "eightyTo99"="grey60", "knownInfection"="black")
```

```{r}
fig5a = ggplot(data=prediction, aes(x, y, color=week35), legend=FALSE) + 
  geom_point() + 
  scale_colour_manual(values = thecol, guide=FALSE) +
  theme_bw() +
  theme_void() 
ggsave(filename="/Users/nwelch/prelim/report/figures/figure_5a.png", fig5a, 
       dpi=300, height=6, width=3.5)
fig5a
```

```{r}
fig5b = ggplot(data=prediction, aes(x, y, color=week40), legend=FALSE) + 
  geom_point() + 
  scale_colour_manual(values = thecol, guide=FALSE) +
  theme_bw() +
  theme_void()
ggsave(filename="/Users/nwelch/prelim/report/figures/figure_5b.png", fig5b, 
       dpi=300, height=6, width=3.5)
fig5b
```
```{r}
fig5c = ggplot(data=prediction, aes(x, y, color=week50), legend=FALSE) + 
  geom_point() + 
  scale_colour_manual(values = thecol, guide=FALSE) +
  theme_bw() +
  theme_void() 
ggsave(filename="/Users/nwelch/prelim/report/figures/figure_5c.png", fig5c, 
       dpi=300, height=6, width=3.5)
fig5c
```

```{r, echo=FALSE}
fig5d = ggplot(data=prediction, aes(x, y, color=week60), legend=FALSE) + 
  geom_point() + 
  scale_colour_manual(values = thecol, guide=FALSE) +
  theme_bw() +
  theme_void() 
ggsave(filename="/Users/nwelch/prelim/report/figures/figure_5d.png", fig5d, 
       dpi=300, height=6, width=3.5)
fig5d
```

```{r, warning=FALSE}
g = grid.arrange(fig5a, fig5b, fig5c, fig5d, ncol=4)
ggsave(filename="/Users/nwelch/prelim/report/figures/figure_5.png", g, 
       dpi=300, height=6, width=12)
```

```{r}
n=100
x = seq(0, 30, 30/n)

get_density = function(i){
  smpl = rnorm(n+1)
  pdf = dnorm(smpl)
  tmp = cumsum(pdf)
  cdf = tmp/tail(tmp, 1)
  return(data.frame(i, x, cdf))
}

tmp = lapply(1:100, get_density)
pdfs = plyr::ldply(tmp, data.frame)

p = ggplot(legend=FALSE) + 
  xlab("week") + 
  ylab("density") + 
  theme(axis.line = element_line(color='black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 

fig3a = p + xlim(0,5) + ylim(0,0.25) + 
  geom_path(data=pdfs[pdfs$i==1,], aes(x,cdf)) +
  geom_path(data=pdfs[pdfs$i==2,], aes(x,cdf), color='darkgray') +
  geom_path(data=pdfs[pdfs$i==3,], aes(x,cdf), color='gray') +
  geom_path(data=pdfs[pdfs$i==4,], aes(x,cdf), color='darkgrey', 
            linetype = "longdash") +
  geom_path(data=pdfs[pdfs$i==5,], aes(x,cdf), color='grey', 
            linetype = "longdash") +
  geom_path(data=pdfs[pdfs$i==6,], aes(x,cdf))

fig3b = p + xlim(24,30) + ylim(0.75, 1) + 
  geom_path(data=pdfs[pdfs$i>6,], aes(x,cdf), color='gray') +
  geom_path(data=pdfs[pdfs$i==1,], aes(x,cdf)) +
  geom_path(data=pdfs[pdfs$i==2,], aes(x,cdf), color='darkgray') +
  geom_path(data=pdfs[pdfs$i==3,], aes(x,cdf), color='gray') +
  geom_path(data=pdfs[pdfs$i==4,], aes(x,cdf), color='darkgrey', 
            linetype = "longdash") +
  geom_path(data=pdfs[pdfs$i==5,], aes(x,cdf), color='grey', 
            linetype = "longdash") +
  geom_path(data=pdfs[pdfs$i==6,], aes(x,cdf))

```

```{r, warning=FALSE}
grid.arrange(fig3a, fig3b, ncol=2)
```

```{r}
n=100
x = seq(0, 30, 30/n)

set.seed(1234)

go_walk = function( i=1, n=30, lmda=1, a0=0 ){
  # This creates the random walk and saves the value of the last poistion in Sn.
  X = rpois(n, lambda=lmda)
  S = c(a0, a0 + cumsum( X ))
  sim = rep(i, n+1)
  week = seq(0, n, 1)
  out = data.frame(week, S, sim)
  return(out)
}

tmp = lapply(1:100, go_walk)
simInfection = plyr::ldply(tmp, data.frame)

tmp = lapply(split(simInfection$S, simInfection$week), 
                        function(i) quantile(i, c(0.025, 0.5, 0.975)))
simInfectionCI = plyr::ldply(tmp)
colnames(simInfectionCI) = c('week', 'lower', 'med', 'upper')
simInfectionCI$week  = as.numeric(simInfectionCI$week)

p = ggplot(legend=FALSE) + 
  xlab("week") + 
  ylab("density") + 
  xlim(0,30)+
  theme(axis.line = element_line(color='black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank()) 

fig4a = p + geom_path(data=simInfection, aes(week, S), colour='darkgray') +
  geom_path(data=simInfectionCI, aes(week, lower), color='black') +
  geom_path(data=simInfectionCI, aes(week, med), color='black') +
  geom_path(data=simInfectionCI, aes(week, upper), color='black') +
  geom_vline(xintercept=timeInspected, color='grey', linetype = "longdash")+
  ylab("Infections per Week")

fig4b = p + geom_path(data=simInfection, aes(week, S), colour='darkgray') +
  geom_path(data=simInfectionCI, aes(week, lower), color='black') +
  geom_path(data=simInfectionCI, aes(week, med), color='black') +
  geom_path(data=simInfectionCI, aes(week, upper), color='black')+
  ylab("Infections per Week")

fig4c = p + geom_path(data=simInfection, aes(week, S^(0.534)-week),
                      colour='darkgray') +
  geom_path(data=simInfectionCI, aes(week, lower^(0.534)-week), color='black') +
  geom_path(data=simInfectionCI, aes(week, med^(0.534)-week), color='black') +
  geom_path(data=simInfectionCI, aes(week, upper^(0.534)-week), color='black') +
  geom_vline(xintercept=timeInspected, color='grey', linetype = "longdash")

fig4d = p + geom_path(data=simInfection, aes(week, S^(0.534)-week),
                      colour='darkgray') +
  geom_path(data=simInfectionCI, aes(week, lower^(0.534)-week), color='black') +
  geom_path(data=simInfectionCI, aes(week, med^(0.534)-week), color='black') +
  geom_path(data=simInfectionCI, aes(week, upper^(0.534)-week), color='black')
```

```{r, warning=FALSE}
grid.arrange(fig4a, fig4b, fig4c, fig4d, ncol=2, nrow=2)
```
```{r, warning=FALSE}
g = grid.arrange(fig4a, fig4b, ncol=2)
g
ggsave(filename="/Users/nwelch/prelim/report/figures/figure_4.png", g, 
       dpi=300, height=4, width=12)
```


